---
title: "EDA"
author: "Weiling Li"
date: "4/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message = F,warning = F)
library(tidyverse)
library(magrittr)
library(readxl)
library(lubridate)
source("functions.R")
```

## Load data


### read data and specify correct NA label

```{r read data}
main_record <- read_xlsx("For Masanao Class - ROMS Full Data Set - March 19th, 2019 Upload.xlsx", sheet = "Master Data Set",col_names = T,na = c("NULL","0000-00-00","Null","null","0000-00-01","0000-00-02","unknown"))

payment_info <- read_xlsx("For Masanao Class - ROMS Full Data Set - March 19th, 2019 Upload.xlsx", sheet = "Claim and Payment Info", col_names = T)
```


```{r glance data}
str(main_record)

str(payment_info)

check_na_df(payment_info)
check_na_df(main_record)
```

There are visit id missing, so when combining both dataset, there will be missing data. One possible solution is to take these unmatched rows and see if we could match them using visit date to the payment dataframe. 



### extract relevant variables and check details

```{r check main record}
colnames(main_record)

#delete ROMS ID that contains NA Visit IDs

invalid_roms <- main_record %>% 
  filter(is.na(`Visit ID`)) %>%
  select(`ROMS ID`) %>% pull() %>% unique()
  

raw <- main_record %>% select(`ROMS ID`,
                              Payer,`Payer Category`,Outcome,
                              Age,`Sex (1=male, 2=female)`,
                              `Outcome Score Range`,
                              `Outcome - Is high score good or bad?`,
                              `Body Region`,
                              Surgical,Classification,`Admission Date`,
                              `Admission Outcome Score`,`Admission Pain`,
                              `Injury Date`,`Surgery Date`,Visits,`Length Of Stay (days)`,
                              `Chronic Pain (Yes/No)`,`Start Back Total`,
                              `Start Back Subscore`) %>% 
  filter(!`ROMS ID` %in% invalid_roms) %>% 
  distinct()

visitid_to_rom <- main_record%>%select(`ROMS ID`,`Visit ID`,`Visit Date`) %>% filter(!is.na(`Visit ID`)) %>% distinct()



payment_info <-left_join(payment_info,visitid_to_rom,by = "Visit ID")

check_na_df(visitid_to_rom)

missing_dates<- payment_info %>% filter(is.na(`ROMS ID`)) %>% select(`Service Date`) %>% unique() %>% pull()
View(payment_info %>% filter(is.na(`ROMS ID`))%>%arrange(`Service Date`))
View(main_record %>% filter(is.na(`Visit ID`) & 
                      `Visit Date` %in% missing_dates) %>% arrange(`Visit Date`))

summary(raw$Visits)

hist(raw$Visits,breaks = 30)
```


## EDA on raw

```{r}
colnames(raw)
check_na_df(raw)
ggplot(raw %>% group_by(`Body Region`) %>% summarise(ct = n()), aes(x = `Body Region`, y = ct)) + geom_col()

## body region needs cleaning

raw %<>% mutate(`Body Region` = tolower(trimws(`Body Region`, which = "both")))

ggplot(raw %>% group_by(`Body Region`) %>% summarise(ct = n()), aes(x = `Body Region`, y = ct,label = ct)) + geom_col() + geom_label()

```

balance, hand, elbow, pelvis and wrist has too little cases, we should group them into others. But we should also look into classification. 

```{r}
nrow(raw %>% group_by(Classification) %>% summarise(ct = n()))
summary(raw %>% group_by(Classification) %>% summarise(ct = n()))
```

Classification is heavily skewed. I think we would be better off using just Body Region.

```{r}
ggplot(raw %>% group_by(`Payer`) %>% summarise(ct = n()), aes(x = `Payer`, y = ct,label = ct)) + geom_col() + geom_label()

hist(raw %>% group_by(`Payer`) %>% summarise(ct = n()) %>% select(ct) %>% pull(),breaks = 100)
summary(raw %>% group_by(`Payer`) %>% summarise(ct = n()) %>% select(ct) %>% pull())
```

Over half of the payers occurs less than 4 times; 3 quaters of them are less than 6 times so we probably dont want to categorize them using payer but rathe payer category.

```{r}
ggplot(raw %>% group_by(`Payer Category`) %>% summarise(ct = n()), aes(x = `Payer Category`, y = ct,label = ct)) + geom_col() + geom_label() + theme(axis.text.x = element_text(angle = 30))
```

The payer of the main record is too imbalanced. try the payer category in the payment table

```{r}
roms_payer <- payment_info %>% select(`ROMS ID`,PayerCategory) %>% filter(!is.na(`ROMS ID`) & !is.na(`PayerCategory`)) %>% distinct()

raw_1 <- left_join(raw,roms_payer,by = "ROMS ID")
check_na_df(raw_1)
ggplot(raw_1 %>% group_by(`PayerCategory`) %>% summarise(ct = n()), aes(x = `PayerCategory`, y = ct,label = ct)) + geom_col() + geom_label() + theme(axis.text.x = element_text(angle = 30))
```


There are 44 NA values, what is causing these NAs? right now, the plan is to use 3 cats, Aetna, BCBS, and Others. We should ask client about the category better.


### Check Admission Outcome

```{r}
ggplot(raw_1 %>% group_by(Outcome) %>% summarise(ct = n()),
       aes(x = Outcome, y = ct, label = ct)) + 
  geom_col() + geom_label() + 
  theme(axis.text.x = element_text(angle = 30))

#There are typos

raw_1 %<>% mutate(Outcome = trimws(tolower(Outcome)))

ggplot(raw_1 %>% group_by(Outcome) %>% summarise(ct = n()),
       aes(x = Outcome, y = ct, label = ct)) + 
  geom_col() + geom_label() + 
  theme(axis.text.x = element_text(angle = 30))
```

After cleaning, there are only 5 cats and they all have sufficient records.

Examine Age
```{r}
hist(raw_1$Age,breaks = 100)

## there are negative ages
## impute with mean age
## 
median_age <- median(raw_1$Age[raw_1$Age >0])
raw_1 %<>% mutate(Age = if_else(Age < 0,median_age, Age))
hist(raw_1$Age,breaks = 100)
```

Encorporate Haozhe's EDA, with minor tweaks. days is to granule, at least using weeks or even months?

```{r}
raw_2 <- raw_1 %>% mutate( weeksdiff_injury_to_ads = 
           as.numeric(
             difftime(raw_1$`Admission Date` , raw_1$`Injury Date`, units = 'weeks')
             ) ) %>%  # add a column of days from injury to admission
  mutate( weeksdiff_surgery_to_ads = 
           as.numeric(
             difftime(raw_1$`Admission Date` , raw_1$`Surgery Date`, units = 'weeks')
             ) ) %>% # add a column of days from surgery to admission
  mutate( `Sex (1=male, 2=female)` = if_else(`Sex (1=male, 2=female)`==1, 0, 1) ) %>%
  rename_( 'Sex' = '`Sex (1=male, 2=female)`' ) %>% # 0 -- male , 1 -- female
  mutate( `Chronic Pain (Yes/No)` = if_else(`Chronic Pain (Yes/No)`== 'Yes', 1 , 0) ) %>%
  rename_('Chronic_Pain' = '`Chronic Pain (Yes/No)`') %>%  # 0 -- no chronic pain, 1 -- chronic pain
  mutate(`Outcome - Is high score good or bad?` = if_else(`Outcome - Is high score good or bad?`== 'Good', 1, 0) ) %>%
  rename_('Outcome_judge' = '`Outcome - Is high score good or bad?`') # 0 -- high score is bad, 1 -- high score is good
```


split outcome score range to the lower end and upper end. calculate a base admission score from 0 to 1 which 0 indicates bad
```{r}
raw_3 <- raw_2 %>% separate(col = `Outcome Score Range`,into = c("lower_score","upper_score"),sep = "-",convert = T) %>%
  mutate(scale_temp = as.numeric(upper_score) - as.numeric(lower_score),
         admin_score = if_else(Outcome_judge == 1, 
                               round((`Admission Outcome Score`-lower_score)/scale_temp,2),
                               round((upper_score - `Admission Outcome Score`)/scale_temp,2)),
         admin_pain = (10-`Admission Pain`)/10) %>%
  select(-scale_temp,
         -lower_score,
         -upper_score, 
         -Outcome_judge,
         -`Admission Outcome Score`,
         -`Admission Pain`)

hist(raw_3$admin_pain,breaks = 11)

```

calculate admission date to the end of the semester

```{r}
summary(year(raw_3$`Admission Date`))
## min 2016 max 2019
dates <- read_csv("semesterdates.csv") %>% pull()
raw_4 <- raw_3 %>% mutate(weeks_to_smsend = round(if_else(`Admission Date` < dates[1],
                                           as.numeric(difftime( dates[1], raw_3$`Admission Date`, units = 'weeks')),
                                           if_else(`Admission Date` < dates[2],
                                                   as.numeric(difftime( dates[2], raw_3$`Admission Date`, units = 'weeks')),
                                           if_else(`Admission Date` < dates[3],
                                                   as.numeric(difftime( dates[3], raw_3$`Admission Date`, units = 'weeks')),
                                           if_else(`Admission Date` < dates[4],
                                                   as.numeric(difftime( dates[4], raw_3$`Admission Date`, units = 'weeks')),
                                           if_else(`Admission Date` < dates[5],
                                                   as.numeric(difftime( dates[5], raw_3$`Admission Date`, units = 'weeks')),
                                           if_else(`Admission Date` < dates[6],
                                                   as.numeric(difftime( dates[6], raw_3$`Admission Date`, units = 'weeks')),
                                           if_else(`Admission Date` < dates[7],
                                                   as.numeric(difftime( dates[7], raw_3$`Admission Date`, units = 'weeks')),
                                                   as.numeric(difftime( dates[8], raw_3$`Admission Date`, units = 'weeks'))))))))),2)) %>%
  select(-`Admission Date`,
         -`Injury Date`)

```

## Clean all the unnecessary cols

```{r}
raw_5 <- raw_4 %>% select(-Payer,
                          -`Payer Category`,
                          -Classification,
                          -`Surgery Date`,
                          -Visits,
                          -`Length Of Stay (days)`,
                          -`Start Back Total`,
                          -`Start Back Subscore`) %>%
  distinct()
```

## Clean admission date issue

```{r}
temp_1 <- raw_5 %>% 
  select(-weeksdiff_injury_to_ads,-weeksdiff_surgery_to_ads,-weeks_to_smsend,-admin_score,-admin_pain,-Age,-Chronic_Pain) %>% 
  distinct()

temp_3 <- temp_1 %>% 
  group_by(`ROMS ID`,Sex,`Body Region`,Surgical,PayerCategory) %>% 
  summarise(outcome_ct = n()) %>% 
  ungroup()

temp_2 <- raw_5 %>% 
  select(`ROMS ID`,Outcome,Surgical,`Body Region`,Surgical,PayerCategory,Chronic_Pain,Age,
         weeksdiff_injury_to_ads,weeksdiff_surgery_to_ads,weeks_to_smsend,admin_score,admin_pain) %>%
  mutate(weeks_to_treat = if_else(is.na(weeksdiff_injury_to_ads),weeksdiff_surgery_to_ads,weeksdiff_injury_to_ads)) %>%
  group_by(`ROMS ID`,Surgical,`Body Region`,Surgical,PayerCategory) %>% summarise(weeks_to_treat = min(weeks_to_treat),
                                                     weeks_to_smsend = max(weeks_to_smsend),
                                                     admin_pain = min(admin_pain),
                                                     admin_score = min(admin_score),
                                                     Age = round(max(Age)),
                                                     Chronic_Pain = max(Chronic_Pain)) %>%
  ungroup()

check_na_df(temp_2)
check_na_df(temp_1)

raw_cleaned <- left_join(temp_3,temp_2) %>% rename_('body_region' = "`Body Region`")

```

## Assign new Category

```{r}
payercat <- c("Blue Cross Blue Shield","Aetna","Medicare")
bodyregion <- c("cervical","foot/ankle","hip","knee","lumbar","shoulder","thoracic")
raw_cleaned_1<- raw_cleaned %>% mutate(PayerCategory = if_else(PayerCategory %in% payercat,PayerCategory,"Other"),
                       body_region = if_else(body_region %in% bodyregion,body_region,"other"))
check_na_df(raw_cleaned_1)
```

There are 3 weeks_to_treat missing, impute using mean

```{r}
mean_week <- mean(raw_cleaned_1$weeks_to_treat,na.rm = T)
raw_cleaned_1[is.na(raw_cleaned_1$weeks_to_treat), ] <- mean_week
check_na_df(raw_cleaned_1)
```

Right now, only the number of survey given by the pt is considered as imput

## Add payment info

```{r}
payment_ROM <- payment_info %>% 
  filter(!is.na(`ROMS ID`)) %>%
  group_by(`ROMS ID`) %>% summarise(total = sum(Amount),
                                    expected_tot = sum(Expected))

check_na_df(payment_ROM)
```

## Combine the two dataframe

```{r}
md_data <- left_join(raw_cleaned_1,payment_ROM) %>% filter(!is.na(total)) ## exclude ones that dont have a total

check_na_df(md_data)
```

